---
title: "FA 7"
author: "PATAYON, SPIKE LEE-ROY V"
date: "2024-02-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

with λ =4 per minute, use R to obtain:
(a) P(T < or equal to 0.25) = P(time between submissions is at most 15 seconds);
(b) P(T > 0.5) = P(time between submissions is greater than 30 seconds);
(c) P(0.25 < T < 1) = P(time between submissions is between 15 second
and 1 minute)

```{r}
# Rate parameter
lambda <- 4

# a
prob_a <- pexp(0.25, rate = lambda)

# b
prob_b <- 1 - pexp(0.5, rate = lambda)

# c
prob_c <- pexp(1, rate = lambda) - pexp(0.25, rate = lambda)

# results
cat("P(T <= 0.25) =", prob_a, "\n")
cat("P(T > 0.5) =", prob_b, "\n")
cat("P(0.25 < T < 1) =", prob_c, "\n")
```
  
The average rate of job submissions in a computer center is 2 per minute. If it can be assumed that the number of submissions per minute has a Poisson distribution, calculate the probability that:

(a) more than two jobs will arrive in a minute;
(b) at least 30 seconds will elapse between any two jobs;
(c) less than 30 seconds will elapse between jobs;
(d) a job will arrive in the next 30 seconds, if no jobs have arrived in the last
30 seconds.

```{r}
# rate of job submissions or the lambda
lambda <- 2

# a
prob_more_than_two <- 1 - ppois(2, lambda)

# b
time_between_jobs <- 30/60
prob_at_least_30_seconds <- ppois(0, lambda * time_between_jobs)

# c
prob_less_than_30_seconds <- 1 - prob_at_least_30_seconds

# d
prob_job_next_30_seconds <- 1 - ppois(0, lambda * time_between_jobs)

#results
cat("Probability of more than two jobs arriving in a minute:", prob_more_than_two, "\n")
cat("Probability of at least 30 seconds between any two jobs:", prob_at_least_30_seconds, "\n")
cat("Probability of less than 30 seconds between jobs:", prob_less_than_30_seconds, "\n")
cat("Probability of a job arriving in the next 30 seconds:", prob_job_next_30_seconds, "\n")

```
A website receives an average of 15 visits per hour, which arrive following a
Poisson distribution:

(a) Calculate the probability that at least 10 minutes will elapse without
a visit
(b) What is the probability that in any hour, there will be less than eight visits?
(c) Suppose that 15 minutes have elapsed since the last visit, what is the probability that a visit will occur in the next 15 minutes
(d) Calculate the top quartile, and explain what it means

```{r}
# Average visits per hour
lambda <- 15

# a
time_without_visit <- 10 / 60
prob_no_visit <- ppois(0, lambda * time_without_visit, lower.tail = TRUE)
prob_at_least_one_visit <- 1 - prob_no_visit

# b
prob_less_than_eight_visits <- ppois(7, lambda, lower.tail = TRUE)

# c
time_since_last_visit <- 15 / 60
prob_visit_in_next_15_min <- 1 - ppois(0, lambda * time_since_last_visit, lower.tail = TRUE)

# d
top_quartile <- qpois(0.75, lambda)

prob_at_least_one_visit = round(prob_at_least_one_visit, digits=2)
prob_less_than_eight_visits = round(prob_less_than_eight_visits, digits=2)
prob_visit_in_next_15_min  = round(prob_visit_in_next_15_min, digits=2)
# result
print(paste("Probability of at least 10 minutes without a visit:", prob_at_least_one_visit))
print(paste("Probability of less than eight visits in any hour:", prob_less_than_eight_visits))
print(paste("Probability of a visit occurring in the next 15 minutes:", prob_visit_in_next_15_min))
print(paste("Top quartile (75th percentile):", top_quartile))
```


